Sulfur Dioxide on Water's Surface

### 2011-06-07 EquilibratedOrientation.tex
Draft 5

Geometric analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules at different depths from the water surface. At each distance from the surface location, an orientation profile was created for both the \wat~and \suldiox~molecules. The orientation distribution for the angles $\theta$ and $\phi$ at all depths were combined to form 2D intensity plots show how the molecular orientation distributions change with distance to the surface location. These plots allow for a visual interpretation of how the net orientations are affected when moving from the gas phase to the surface, and then further into the aqueous bulk. Both the neat-water, with only a single \suldiox~introduced, and the high-concentration saturated system were analyzed. In the case of the neat-water system, the introduction of a single \suldiox~does not greatly affect the water surface, and the results of the water orientation are very similar to a neat-water system without any adsorbed solutes.

@@

The plots of $\cos(\phi)$ are both isotropic, although the neat-water system plot is quite noisy from the small data set. Because the \suldiox~near the surface is oriented perpendicularly to the interface, the $\cos(\phi)$ orientation is expected to be isotropic. Further from the water surface where the bisector orientation becomes isotropic, the $\cos(\phi)$ distribution also becomes isotropic. For the \suldiox~orientation, the $\phi$ angle does not provide further information regarding the surface behavior.

The plots indicate that \suldiox~orients similarly at both the low and high concentration water surface environments. The plots of $\cos(\phi)$ for both concentrations show a isotropic behavior. However, the $\theta$ values follow similar trends at both concentrations indicating that the \suldiox~sulfur orients down towards the water phase, with both oxygens pointing away from the surface once the \suldiox~is within approx. 5\angs~of the surface. This behavior continues down to at least 5\angs~below the surface, notwithstanding any chemical reactions that may occur that are not simulated using the classical MD techniques utilized in this work.

>>>>

Geometric analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules at different depths from the water surface. At each distance from the surface location, an orientation profile was created for both the \wat~and \suldiox~molecules. The orientation distribution for the angles $\theta$ and $\phi$ at all depths were combined to form 2D intensity plots that show how the molecular orientation distributions change with distance to the surface location. These plots allow for a visual interpretation of how the net orientations are affected when moving from the gas phase to the surface, and then further into the aqueous bulk. Both the neat-water, with only a single \suldiox~introduced, and the high-concentration saturated system were analyzed. In the case of the neat-water system, the introduction of a single \suldiox~does not greatly affect the water surface, and the results of the water orientation are very similar to a neat-water system without any adsorbed solutes (not shown).

@@

The plots of $\phi$ are both isotropic, although the neat-water system plot is quite noisy from the small data set. Because the \suldiox~near the surface is oriented perpendicularly to the interface, the $\cos(\phi)$ orientation is expected to be isotropic. Further from the water surface where the bisector orientation becomes isotropic, the $\cos(\phi)$ distribution also becomes isotropic. For the \suldiox~orientation, the $\phi$ angle does not provide further information regarding the surface behavior.

The plots indicate that \suldiox~orients similarly at both the low and high concentration water surface environments. The plots of $\phi$ for both concentrations exhibit isotropic distributions. However, the $\theta$ values follow similar trends at both concentrations indicating that the \suldiox~sulfur orients down towards the water phase, with both oxygens pointing away from the surface once the \suldiox~is within approx. 5\angs~of the surface. This behavior continues down to at least 5\angs~below the surface, notwithstanding any chemical reactions that may occur that are not simulated using the classical MD techniques utilized in this work.

### 2011-06-07 Methods.tex
Draft 5

Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation. Figure \ref{fig:density-flaw}a shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work using the original uncorrected coordinates of the system. The water density profile and location (dashed line in figure \ref{fig:density-flaw}a) is produced by averaging the instantaneous density profile at each timestep in the simulation over all the timesteps. The water profile was then fit to a $\tanh$ function (black line) to extract the position and width parameters of the water surface. The fitted Gibbs dividing surface has a width of 3.77\angs, which is comparable to values reported for similar neat-water systems.\cite{Dang1997,Hore2008} A bulge in the gas-phase ($\leq 52\AA$) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory. Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.

>>>>

Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation. Figure \ref{fig:density-flaw}a shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work using the original uncorrected coordinates of the system. The water density profile and location (dashed line in figure \ref{fig:density-flaw}a) is produced by averaging the instantaneous density profile at each timestep in the simulation over all the timesteps. The water profile was then fit to a $\tanh$ function (black line) to extract the position and width parameters of the water surface. The fitted Gibbs dividing surface has a width of 3.77\angs, which is comparable to values reported for similar neat-water systems.\cite{Dang1997,Hore2008} A bulge in the gas-phase ($\leq 52$\angs) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory. Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.

### 2011-06-07 TransitOrientation.tex
Draft 5

	The orientation of \suldiox~molecules throughout the aqueous adsorption process was monitored during the transit SMD simulations. The angles $\theta$ and $\phi$ (figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the transiting \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems. The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface, resulting in the 2-dimensional angle and depth-profile histograms shown in figure \ref{fig:so2-transit-angles}. Only the surface into which the transiting \suldiox~was steered is used in the analyses. In all the data sets, the populations of the angle histograms at each distance from the surface were normalized to aid in comparison of regions with differing \suldiox~residence times.

@@

	From its starting position 20\angs~above the water surface, until the \suldiox~moves to within 10\angs~of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$. This can be seen in the plots as mostly uniform coloration at a given distance from the surface independent of $\cos$ of $\phi$. Near the surface the bisector angle $\theta$ becomes more perpendicular ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase. At the point when the \suldiox~reaches the surface location (distance $=0$\angs), the bisector is perpendicular in both the neat-water and saturated systems. Figure \ref{fig:so2-surface-angles}a depicts \suldiox~near the water surface with the bisector aligned perpendicular to the interface. In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface. The main difference between the two systems is where the point of the transition from isotropic to preferred orientations is found.

  % talk about theta differences
  Looking more closely at the approach, the $\cos(\theta)$ distribution above 10\angs~from the surface is isotropic. For the neat-water surface, a transition occurs at approximately 5\angs~from the surface where the \suldiox~orients with its bisector perpendicular to the water surface. This coincides with the approximate distance of bonds formed to the outer-most water molecules. Below 5\angs, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly. In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approximately 8\angs~above the surface. The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule. Also, topmost water molecules from the surface move up to a few \angs~inside the adsorbed \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system. It is remarkable that the orientational trend appears so strong in the $\cos(\theta)$ plots even with so little data as was collected from the single \suldiox~molecule of each simulation. From onset of orientation above the surface until 10\angs~below, the \suldiox~holds a preferred orientation.

>>>>

	The orientation of \suldiox~molecules throughout the aqueous adsorption process was monitored during the transit SMD simulations. The angles $\theta$ and $\phi$ of the transiting \suldiox~(figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems. The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface, resulting in the 2-dimensional angle and depth-profile histograms shown in figure \ref{fig:so2-transit-angles}. Only the surface into which the transiting \suldiox~was steered is used in the analyses. In all the data sets, the populations of the angle histograms at each distance from the surface were normalized to aid in comparison of regions with differing \suldiox~residence times.

@@

	From its starting position 20\angs~above the water surface, until the \suldiox~moves to within 10\angs~of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$. This can be seen in the plots as mostly uniform coloration at a given distance from the surface independent of $\cos$ of $\phi$. Near the surface the bisector angle $\theta$ becomes more perpendicular to the water surface ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase. At the point when the \suldiox~reaches the surface location (distance $=0$\angs), the bisector is perpendicular in both the neat-water and saturated systems. Figure \ref{fig:so2-surface-angles}a depicts \suldiox~near the water surface with the bisector aligned perpendicular to the interface. In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface. The main difference between the two systems is where the point of the transition from isotropic to preferred orientations is found.

  % talk about theta differences
  Looking more closely at the approach, the $\theta$ distribution above 10\angs~from the surface is isotropic. For the neat-water surface, a transition occurs at approximately 5\angs~from the surface where the \suldiox~orients with its bisector perpendicular to the water surface. This coincides with the approximate distance of bonds formed to the outer-most water molecules. Below 5\angs, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly. In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approximately 8\angs~above the surface. The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule. Also, topmost water molecules from the surface move up to a few \angs~inside the adsorbed \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system. It is remarkable that the orientational trend appears so strongly in the $\theta$ plots even with so little data as was collected from the single \suldiox~molecule of each simulation. From onset of orientation above the surface until 10\angs~below, the \suldiox~holds a preferred orientation.
